### 2023-09-05 proofs/main.tex
WIP: separate the algorithmic proofs
Grammar update: Γ = set
Rearrange proofs

\newcommand{\depth}[1]{\mathsf{depth}(#1)}
\newcommand{\size}[1]{\mathsf{size}(#1)}

@@

\section{Declarative System}
First, we present the top-level system, which is easy to understand.

@@

\ottgrammartabular{
  \ottiP\ottinterrule
  \ottiN\ottinterrule
  \ottv\ottinterrule
  \ottc\ottinterrule
}

@@

\subsection{Contexts}

\begin{definition}[Declarative Type Context]
  \hfill \\
  Declarative type context $[[Γ]]$ is represented by a set of 
  type variables. The concatenation $[[Γ1, Γ2]]$ means the 
  union of two contexts  $[[{Γ1} ∪ {Γ2}]]$.
\end{definition}

@@

\begin{definition}[Substitution Signature]
  The signature $[[Γ' ⊢ σ : Γ]]$ means that
  \begin{enumerate}
    \item for any $[[α± ∊ {Γ}]], [[ Γ' ⊢ [σ]α± ]]$; and
    \item for any $[[α± ∉ {Γ'}]], [[ [σ]α± = α± ]]$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}[Monadic Substitution Composition]
  Suppose that $[[Γ' ⊢ σ1 : Γ]]$.
  Then we define $[[σ2 <=< σ1]]$ as $[[(σ2 ○ σ1)|{Γ}]]$.
\end{definition}
Notice that the result of $[[σ2 <=< σ1]]$ depends on the 
specification of $[[σ1]]$, which is not unique. 
However, we assume that the used specification clear from the 
context of the proof.

\begin{definition}[Equivalent Substitutions]
  The substitution equivalence judgement $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$ 
  indicates that on the domain $[[Γ]]$, 
  the result of $[[σ1]]$ and $[[σ2]]$ are equivalent in context $[[Γ']]$.
  Formally, for any $[[α± ∊ {Γ}]], [[ Γ' ⊢ [σ1]α± ≈ [σ2]α± ]]$.
\end{definition}

@@

\subsection{Declarative Typing}
\ottdefnsDT

@@

\subsection{Well-Formedness}
\ottdefnsWFT
\ottdefnsWFAT

\section{Algorithm}

@@

In several places of our algorithm, 
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.

\begin{definition}[Variable Algorithmization]
  Suppose that $[[nas]]$ is a list of negative type variables
  and $[[nuas]]$ is a list of negative algorithmic variables of the same length. 
  Then $[[ nuas/nas ]]$ is a substitution turning each $[[αi⁻]]$ from $[[nas]]$ into 
  $[[αî⁻]]$ from $[[nuas]]$.
  Positive algorithmization $[[ puas/pas ]]$
  is defined symmetrically.
\end{definition}

\subsection{Contexts}

\begin{definition}[Algorithmic Type Context]
  \hfill \\
  Algorithmic type context $[[Ξ]]$ is represented by a set of 
  \emph{algorithmic} type variables ($[[α̂⁺]]$, $[[α̂⁻]]$, $[[β̂⁺]]$, \dots).
\end{definition}

@@

@@

% \subsection{Algorithmic Equivalence}
% \ottdefnsEOneA

@@

\subsection{Typing}
\ottdefnsAT

\section{Proofs}

\subsection{Type well-formedness}
\input{../_build/_gen/tex/wf-lemmas.tex}

@@

\subsection{Overview}
\input{../_build/_gen/tex/lemmas-overview.tex}

@@

@@

@@

\subsection{Declarative Typing}
\input{../_build/_gen/tex/decl-typing-lemmas.tex}

\subsection{Algorithmic Typing}
\input{../_build/_gen/tex/alg-typing-lemmas.tex}

>>>>

\newcommand{\depth}[1]{\ensuremath{\mathsf{depth}(#1)}}
\newcommand{\size}[1]{\ensuremath{\mathsf{size}(#1)}}

@@

\section{Declarative Type Systems}

@@

\ottgrammartabular{
  \ottiP\ottinterrule
  \ottiN\ottinterrule
}

@@

\subsection{Contexts and Well-formedness}

\begin{definition}[Declarative Type Context]
  \hfill \\
  Declarative type context $[[Γ]]$ is represented by a set of 
  type variables. The concatenation $[[Γ1, Γ2]]$ means the 
  union of two contexts $[[Γ1 ∪ Γ2]]$.
\end{definition}

$[[Γ ⊢ iP]]$ and $[[Γ ⊢ iN]]$ denote that the type is well-formed in the context $[[Γ]]$,
which, in fact, means that each free type variable of the type is contained in $[[Γ]]$
(it will be shown later in \cref{lemma:wf-soundness,lemma:wf-ctxt-equiv}).

\hfill\\
\ottdefnsWFT

@@

\begin{definition}[Substitution Signature]
  The signature $[[Γ' ⊢ σ : Γ]]$ means that
  \begin{enumerate}
    \item for any $[[α± ∊ Γ]], [[ Γ' ⊢ [σ]α± ]]$; and
    \item for any $[[α± ∉ Γ']], [[ [σ]α± = α± ]]$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}[Monadic Substitution Composition]
  Suppose that $[[Γ' ⊢ σ1 : Γ]]$.
  Then we define $[[σ2 <=< σ1]]$ as $[[(σ2 ○ σ1)|Γ]]$.
\end{definition}
Notice that the result of $[[σ2 <=< σ1]]$ depends on the 
specification of $[[σ1]]$, which is not unique. 
However, we assume that the used specification clear from the 
context of the proof.

\begin{definition}[Equivalent Substitutions]
  The substitution equivalence judgement $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$ 
  indicates that on the domain $[[Γ]]$, 
  the result of $[[σ1]]$ and $[[σ2]]$ are equivalent in context $[[Γ']]$.
  Formally, for any $[[α± ∊ Γ]], [[ Γ' ⊢ [σ1]α± ≈ [σ2]α± ]]$.
\end{definition}

@@

@@

\section{Algorithmic Type System}

@@

In several places of our algorithm, in particular, during anti-unification,
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.

\begin{definition}[Variable Algorithmization]
  Suppose that $[[nas]]$ is a list of negative type variables
  and $[[nuas]]$ is a list of negative algorithmic variables of the same length. 
  Then $[[ nuas/nas ]]$ is a substitution-like procedure turning each $[[αi⁻ ∊ {nas}]]$ 
  into $[[αî⁻ ∊ {nuas}]]$.
\end{definition}

Conversely, we can turn algorithmic type variables
into declarative type variables via \emph{dealgorithmization}.

\begin{definition}[Variable Dealgorithmization]
  Suppose that $[[nuas]]$ is a list of negative algorithmic variables
  and $[[nas]]$ is a list of negative type variables of the same length. 
  Then $[[ nas/nuas ]]$ is a substitution-like procedure turning each
  $[[αî⁻ ∊ {nuas}]]$ into $[[αi⁻ ∊ {nas}]]$.
\end{definition}

\subsection{Contexts and Well-formedness}

\begin{definition}[Algorithmic Type Context]
  \hfill \\
  Algorithmic type context $[[Ξ]]$ is represented by a set of 
  \emph{algorithmic} type variables ($[[α̂⁺]]$, $[[α̂⁻]]$, $[[β̂⁺]]$, \dots).
  The concatenation $[[Ξ1, Ξ2]]$ means the union of two contexts $[[Ξ1 ∪ Ξ2]]$.
\end{definition}

@@

\hfill\\
\ottdefnsWFAT

@@

@@

\section{Declarative Typing}

\bigskip
\ottgrammartabular{
  \ottv\ottinterrule
  \ottc\ottinterrule
}

\subsection{Declarative Type Inference}
\ottdefnsDT

\section{Algorithmic Typing}

\subsection{Algorithmic Type Inference}
\ottdefnsAT

\section{Properties of the Declarative Type System}

\subsection{Type Well-formedness}
\input{../_build/_gen/tex/wf-lemmas.tex}

@@

@@

\section{Properties of the Algorithmic Type System}

\subsection{Algorithmic Type Well-formedness}
\input{../_build/_gen/tex/wf-algo-lemmas.tex}

\subsection{Substitution}
\input{../_build/_gen/tex/subst-algo-lemmas.tex}

\subsection{Normalization}
\input{../_build/_gen/tex/norm-alg-lemmas.tex}

\subsection{Equivalence}
\input{../_build/_gen/tex/equiv-alg-lemmas.tex}

@@

\section{Properties of the Declarative Typing}
\input{../_build/_gen/tex/decl-typing-lemmas.tex}

\section{Properties of the Algorithmic Typing}

@@

\subsection{Correctness of the Typing Algorithm}
\input{../_build/_gen/tex/alg-typing-lemmas.tex}

### 2023-09-05 proofs/tex/alg-pos-subtyping-lemmas.tex
Rearrange proofs

            By \cref{lemma:quant-rule-decomposition} applied twice, 
            $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ ∃ncs.iQ0 ]]$ implies
            $[[ Γ,ncs  ⊢ [iNs/nbs][uσ]α̂⁺ ≥ iQ0 ]]$ for some $[[iN]]$, 
            and since $[[ {nbs} ∩ fv([uσ]α̂⁺) ⊆ {nbs} ∩ {Θ(α̂⁺)} ⊆ {nbs} ∩ Γ = ∅ ]]$,
            $[[ [iNs/nbs][uσ]α̂⁺ = [uσ]α̂⁺ ]]$, that is $[[ Γ,ncs ⊢ [uσ]α̂⁺ ≥ iQ0]]$.

@@

            Now, let us show that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
            To do that, we notice that $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]]$:
            let us consider an arbitrary variable appearing freely in $[[uP0]]$:
            \begin{enumerate}
                \item if this variable is a metavariable $[[αî⁻]] \in [[â⁻*]]$, then
                $[[ [uσ0]αî⁻ ]] = [[ [σ]αi⁻ ]]$ and 
                $[[ [uσ][σ][nas/â⁻*]αî⁻ ]] = [[ [uσ][σ]αi⁻ ]] = [[ [σ]αi⁻ ]]$,
                \item if this variable is a metavariable $[[β̂±]] \in [[ uv(uP0) \ {â⁻*} ]] = [[uv(uP')]]$, then
                $[[ [uσ0]β̂± ]] = [[ [uσ]β̂± ]]$ and $[[ [uσ][σ][nas/â⁻*]β̂± ]] = [[ [uσ][σ]β̂± ]] = [[ [uσ]β̂± ]]$,
                \item if this variable is a regular variable from $[[fv(uP0)]]$, both substitutions do not change it:
                $[[ uσ0 ]]$, $[[ uσ ]]$ and $[[ nas / â⁻* ]]$ act on metavariables, 
                and $[[σ]]$ is defined on $[[nas]]$, however, $[[{nas} ∩ fv(uP0) = ∅]]$.
            \end{enumerate}
            This way, $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]] = [[ [uσ][σ]uP' ]]$,
            and thus, $[[ Γ, nbs ⊢ [uσ0]uP0 ≥ iQ' ]]$.
    \end{caseof}
\end{proof}

>>>>

            By \cref{lemma:quant-rule-decomposition} applied twice, 
            $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ ∃ncs.iQ0 ]]$ implies
            $[[ Γ,ncs  ⊢ [iNs/nbs][uσ]α̂⁺ ≥ iQ0 ]]$ for some $[[iN]]$, 
            and since $[[ {nbs} ∩ fv([uσ]α̂⁺) ⊆ {nbs} ∩ {Θ(α̂⁺)} ⊆ {nbs} ∩ Γ]] = [[ ∅ ]]$,
            $[[ [iNs/nbs][uσ]α̂⁺ = [uσ]α̂⁺ ]]$, that is $[[ Γ,ncs ⊢ [uσ]α̂⁺ ≥ iQ0]]$.

@@

            Now, let us show that $[[Γ, nbs ⊢ [us0]uP0 ≥ iQ]]$.
            To do that, we notice that $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]]$:
            let us consider an arbitrary variable appearing freely in $[[uP0]]$:
            \begin{enumerate}
                \item if this variable is a algorithmic variable $[[αî⁻]] \in [[â⁻*]]$, then
                $[[ [uσ0]αî⁻ ]] = [[ [σ]αi⁻ ]]$ and 
                $[[ [uσ][σ][nas/â⁻*]αî⁻ ]] = [[ [uσ][σ]αi⁻ ]] = [[ [σ]αi⁻ ]]$,
                \item if this variable is a algorithmic variable $[[β̂±]] \in [[ uv(uP0) \ {â⁻*} ]] = [[uv(uP')]]$, then
                $[[ [uσ0]β̂± ]] = [[ [uσ]β̂± ]]$ and $[[ [uσ][σ][nas/â⁻*]β̂± ]] = [[ [uσ][σ]β̂± ]] = [[ [uσ]β̂± ]]$,
                \item if this variable is a regular variable from $[[fv(uP0)]]$, both substitutions do not change it:
                $[[ uσ0 ]]$, $[[ uσ ]]$ and $[[ nas / â⁻* ]]$ act on algorithmic variables, 
                and $[[σ]]$ is defined on $[[nas]]$, however, $[[{nas} ∩ fv(uP0) = ∅]]$.
            \end{enumerate}
            This way, $[[ [uσ0]uP0 ]] = [[ [uσ][σ][nas/â⁻*]uP0 ]] = [[ [uσ][σ]uP' ]]$,
            and thus, $[[ Γ, nbs ⊢ [uσ0]uP0 ≥ iQ' ]]$.
    \end{caseof}
\end{proof}

### 2023-09-05 proofs/tex/anti-unification-lemmas.tex
Rearrange proofs

                This way, by the induction hypothesis:
                \begin{enumerate}
                    \item $[[Γ ; Ξ ⊢ uQ]]$ and $[[Γ ; Ξ' ⊢ uM]]$. 
                    Then by weakening (\cref{todo}), $[[Γ ; Ξ ∪ Ξ' ⊢ uQ]]$ and 
                    $[[Γ ; Ξ ∪ Ξ' ⊢ uM]]$, which implies $[[Γ ; Ξ ∪ Ξ' ⊢ uQ → uM]]$;

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:anti-unification-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions , and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a metavariable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:anti-unification-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:anti-unification-deterministic}).
    \end{itemize}

    Now, we can assume that $[[uM']]$ is not a metavariable. 
    We prove by induction on the derivation of $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on the derivation of $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.

    Since $[[uM']]$ is not a metavariable, 
    the substitution acting on $[[uM']]$ preserves its outer constructor. 
    In other words, 
    $[[ [aus'i] uM' = iNi ]]$ means that $[[uM']]$, 
    $[[iN1]]$ and $[[iN2]]$ have the same outer constructor. 
    Let us consider the algorithmic anti-unification rule corresponding to this constructor, 
    and show that it was successfully applied to anti-unify $[[iN1]]$ and $[[iN2]]$ 
    (or $[[iP1]]$ and $[[iP2]]$).

@@

        Then $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, 
        and $[[aus1]] = [[aus2]] = [[·]]$.
        Since $[[ [aus'i] uM' = iNi ]] = [[α⁻]]$
        and $[[uM']]$ is not a metavariable, $[[uM']] = [[α⁻]]$.
        Then we can take $[[ausr]] = [[·]]$, which satisfies the required properties:
        \begin{itemize}
            \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds vacuously since 
            $[[Ξ' | uv uM']] = [[∅]]$; 
            \item $[[ [ausr] uM' = uM ]]$, that is $[[ [·] α⁻ = α⁻ ]]$
            holds by substitution properties;
            \item the unique determination of $[[ [ausr]α̂⁻]]$ for $[[α̂⁻]] \in [[Ξ' | uv uM']] = [[∅]]$ holds vacuously.
        \end{itemize}

@@

        Then since $[[ [aus'i] uM' = iNi ]] = [[↑iPi]]$ and $[[uM']]$ is not a metavariable,
        $[[uM']] = [[↑uQ']]$, where $[[ [aus'i] uQ' = iPi ]]$. 
        Let us show that $[[(Ξ', uQ', aus'1, aus'2)]]$ 
        is an anti-unifier of $[[iP1]]$ and $[[iP2]]$.
        \begin{enumerate}
            \item $[[Γ ; Ξ' ⊢ uQ']]$ holds by inversion of $[[Γ ; Ξ' ⊢ ↑uQ']]$;
            \item $[[Γ ; · ⊢ aus'i : Ξ']]$ holds by assumption;
            \item $[[ [aus'i] uQ' = iPi ]]$ holds by assumption.
        \end{enumerate}

@@

        Then since $[[ [aus'i] uM' = iNi ]] = [[iPi → iN'i]]$ and $[[uM']]$ is not a metavariable,
        $[[uM']] = [[uQ' → uM'']]$, where $[[ [aus'i] uQ' = iPi ]]$ and $[[ [aus'i] uM'' = iN''i ]]$.

>>>>

                This way, by the induction hypothesis:
                \begin{enumerate}
                    \item $[[Γ ; Ξ ⊢ uQ]]$ and $[[Γ ; Ξ' ⊢ uM]]$. 
                    Then by weakening (\cref{lemma:wf-weakening-algo}), 
                    $[[Γ ; Ξ ∪ Ξ' ⊢ uQ]]$ and 
                    $[[Γ ; Ξ ∪ Ξ' ⊢ uM]]$, which implies $[[Γ ; Ξ ∪ Ξ' ⊢ uQ → uM]]$;

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:anti-unification-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions , and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a algorithmic variable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:anti-unification-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:anti-unification-deterministic}).
    \end{itemize}

    Now, we can assume that $[[uM']]$ is not a algorithmic variable. 
    We prove by induction on the derivation of $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on the derivation of $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.

    Since $[[uM']]$ is not a algorithmic variable, 
    the substitution acting on $[[uM']]$ preserves its outer constructor. 
    In other words, 
    $[[ [aus'i] uM' = iNi ]]$ means that $[[uM']]$, 
    $[[iN1]]$ and $[[iN2]]$ have the same outer constructor. 
    Let us consider the algorithmic anti-unification rule corresponding to this constructor, 
    and show that it was successfully applied to anti-unify $[[iN1]]$ and $[[iN2]]$ 
    (or $[[iP1]]$ and $[[iP2]]$).

@@

        Then $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, 
        and $[[aus1]] = [[aus2]] = [[·]]$.
        Since $[[ [aus'i] uM' = iNi ]] = [[α⁻]]$
        and $[[uM']]$ is not a algorithmic variable, $[[uM']] = [[α⁻]]$.
        Then we can take $[[ausr]] = [[·]]$, which satisfies the required properties:
        \begin{itemize}
            \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds vacuously since 
            $[[Ξ' | uv uM']] = [[∅]]$; 
            \item $[[ [ausr] uM' = uM ]]$, that is $[[ [·] α⁻ = α⁻ ]]$
            holds by substitution properties;
            \item the unique determination of $[[ [ausr]α̂⁻]]$ for $[[α̂⁻]] \in [[Ξ' | uv uM']] = [[∅]]$ holds vacuously.
        \end{itemize}

@@

        Then since $[[ [aus'i] uM' = iNi ]] = [[↑iPi]]$ and $[[uM']]$ is not a algorithmic variable,
        $[[uM']] = [[↑uQ']]$, where $[[ [aus'i] uQ' = iPi ]]$. 
        Let us show that $[[(Ξ', uQ', aus'1, aus'2)]]$ 
        is an anti-unifier of $[[iP1]]$ and $[[iP2]]$.
        \begin{enumerate}
            \item $[[Γ ; Ξ' ⊢ uQ']]$ holds by inversion of $[[Γ ; Ξ' ⊢ ↑uQ']]$;
            \item $[[Γ ; · ⊢ aus'i : Ξ']]$ holds by assumption;
            \item $[[ [aus'i] uQ' = iPi ]]$ holds by assumption.
        \end{enumerate}

@@

        Then since $[[ [aus'i] uM' = iNi ]] = [[iPi → iN'i]]$ and $[[uM']]$ is not a algorithmic variable,
        $[[uM']] = [[uQ' → uM'']]$, where $[[ [aus'i] uQ' = iPi ]]$ and $[[ [aus'i] uM'' = iN''i ]]$.
